---
title: "Amherst History Report"
author: "XX GROUP NAMES HERE XX"
date: "XX DATE 2020"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
    toc: true
---


```{r, setup, include=FALSE}
library(mosaic)  
library(tidyverse)
library(mdsr)
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

\newpage

# Executive Summary (for Biddy)

This should be self-contained and up to one page in length (imagine that is *all* that she would read).

\newpage

# Wrangling

```{r}
loc <- "chapter_files/"
list.files(loc)
sample_chapter <- readLines(paste(loc, "chapter01.txt", sep = ""))
head(sample_chapter, 30)
```

Your wrangling should go here.  If you need to do extensive processing of the chapters then please read in that output here and reference where I can find the other code.

Details matter: please be attentive to aspects of appearance and code style.

# Analysis

This should include your analyses and interpretation. 

# Discussion

Here I am expecting two or three paragraphs

What did you learn?  

What insights did you glean?

# Technical Appendix

In a paragraph or two, describe what wrangling or analysis you did that went above and beyond what was discussed in chapter 19 of MDSR: https://beanumber.github.io/mdsr2e/ch-text.html 
